Dissertation Proposal

### 2023-10-25 paper.tex
clean up some tables
move figures to the paper body

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\textwidth]{images/smol-quizzes-example.png}
  \caption{Screenshot of a SMoL Quiz question.}
  \label{f:smol-quizzes-example}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\textwidth]{images/smol-tutor-example.png}
  \caption{Screenshots of an interpreting question in SMoL Tutor. The
    top-left shows the initial state, where the question is presented
    as an MCQ. If a student chooses a wrong answer, they will receive
    feedback (bottom-left) and will be asked a similar question
    (right). The similar question must be answered by typing.}
  \label{f:smol-tutor-example}
\end{figure}

@@

\begin{table}[t]
  % \begin{tabularx}{250pt}{|c|X|c|X|} \end{tabularx}
  \centering\begin{tabularx}{\linewidth}{
      >{\hsize=.5\hsize\linewidth=\hsize}X
      >{\hsize=1.5\hsize\linewidth=\hsize}X}
    \textbf{Tutorial}                             & \textbf{Goal Sentences}
    \\\hline
    \newterm{scope}: Variable definitions and function definitions &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Referring to an unbound variable leads to an error.
      \item Define "blocks".
      \item SMoL is lexically scoped rather than dynamically scoped.
      \item SMoL disallows defining a variable twice in one block.
      \item SMoL is eager (and evaluates from left to right) rather
            than lazy, reactive, or relational.
    \end{enumerate}
    \\\hline
    \newterm{mut-vars}: Variable updates &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Variable assignment respects the hierarchical structure of blocks.
      \item Variables do not alias.
    \end{enumerate}
    \\\hline
    \newterm{begin}: Sequencing expressions &
    A sequencing expression evaluates its sub-expressions from left to right and
    returns the value of the last sub-expression.
    \\\hline
    \newterm{vectors}: Vectors and vector updates &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Define heap and memory addresses.
      \item Vectors can be referred to by vectors, including themselves.
      \item Vectors are not copied but aliased by bindings.
      \item Constructing vectors doesn't alter environments.
      \item Introducing bindings doesn't alter the (value part of) heap.
    \end{enumerate}
    \\\hline
    \newterm{lambda}: Lambda expressions &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Functions are (first-class) values.
      \item Functions remember the environment where they were created.
      \item Functions can be created with lambda expressions.
      \item A function definition can be viewed as a variable definition plus a lambda expression.
    \end{enumerate}
    \\\hline
    \newterm{local}: Local binding forms &
    Introduce \texttt{let}, \texttt{letrec}, and \texttt{let*}.
    % \\\hline
  \end{tabularx}
  \caption{SMoL Tutor tutorials and their goal sentences.}
  \label{t:tutor-goal-sentences}
\end{table}

@@

  \begin{table}[t]
    \centering
    \begin{tabularx}{\textwidth}{m{0.25\textwidth} l Y}
      \textbf{Misconception}                                                     &
      \textbf{Question}                                                          &
      \textbf{Table of Answers}
      \\\hline
      \miscon{CallByRef} Function calls alias variables.                         &
       % mut-vars.not_aliased_by_funarg_2 (35)
       \lstinputlisting{programs/call-by-ref.txt}                                                                    & {
        \begin{tabularx}{1in}{r l}
          78\%  & \verb|12|      \\
          11\% & \verb|error|      \\
          **10\%   & \verb|0|  \\
          1\%   & \verb|23|  \\
        \end{tabularx}
    }
      \\\hline
      \miscon{CallsCopyStructs} Function calls copy data structures.      &
      % vectors2.alias_with_funcall (48)
      \lstinputlisting{programs/calls-copy-structs.txt}
      & {
        \begin{tabularx}{1in}{r l}
          90\%  & \verb|#(173 0)|      \\
          **10\%   & \verb|#(1 0)|  \\
        \end{tabularx}
    } \\\hline
      \miscon{DeepClosure} Closures copy the \textit{values} of free variables.  &
      % lambda3.lambda_remembers_env (70)
      \lstinputlisting{programs/deep-closure.txt}                                & {
          \begin{tabularx}{1in}{r l}
            86\%  & \verb|4|      \\
            **10\% & \verb|3|      \\
            3\%   & \verb|error|  \\
            1\%   & \verb|lambda| \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{DefByRef} Variable definitions alias variables.                    &
      % mut-vars.not_aliased_by_defvar_1 (29)
      \lstinputlisting{programs/def-by-ref.txt}                                  & {
          \begin{tabularx}{1in}{r Y}
            85\%   & \verb|0 12|                \\
            **12\% & \verb|0 0|                 \\
            2\%    & \verb|error|               \\
            1\%    & depends on implementation. \\
          \end{tabularx}
        }
      \\\hline
      \miscon{DefOrSet}  Both definitions and variable assignments are
      interpreted as follows: if a variable is not defined in the
      current environment, it is defined. Otherwise, it is mutated to
      the new value.
      &
      % program: mut-vars.update_undefined (28)
      \lstinputlisting{programs/def-or-set.txt}                                  & {
          \begin{tabularx}{1in}{r l}
            85\%   & \verb|error| \\
            **15\% & \verb|2|     \\
          \end{tabularx}
        }
      \\\hline
      \miscon{DefsCopyStructs} Variable definitions copy structures recursively. &
      % vectors2.alias_with_defvar (47)
      \lstinputlisting{programs/defs-copy-structs.txt}                                & {
          \begin{tabularx}{1in}{r l}
            **67\% & \verb|#(100)| \\
            30\%   & \verb|#(200)| \\
            1\%    & \verb|#(300)| \\
            1\%    & \verb|error|  \\
          \end{tabularx}
      }                                                                                   \\\hline
    \end{tabularx}
    \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
      with ``**'' represent the misconception. (Part I)}
    \label{t:tutor-grounded-misconceptions-1}
  \end{table}

  \begin{table}[t]

@@

    \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}

@@

          \begin{tabularx}{1in}{r l}

@@

          \begin{tabularx}{1in}{r p{0.2\textwidth}}

@@

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
    \textbf{Misconception}                                                             &
    \textbf{Question}                                                                  &
    \textbf{Table of Answers}                                                              \\\hline
    \surmisedMiscon{NestedDef} Sequences of definitions are
    interpreted as if they are written in nested blocks. A definition
    is not in the scope of later definitions.
    &
    % program: scope.what_is_x_4 (13)
    \lstinputlisting{programs/nested-def.txt}                                          & {
        \begin{tabularx}{1in}{r l}
          92\%  & \verb|2| \\
          **8\% & \verb|1| \\
        \end{tabularx}
    }                                                                                      \\\hline
    \surmisedMiscon{Lazy} Expressions are only evaluated when their values are needed. &
    % order.bad_order (20)
    \lstinputlisting{programs/lazy.txt}                                                & {
        \begin{tabularx}{1in}{r l}
          **57\% & \verb|1 3|   \\
          43\%   & \verb|error| \\
        \end{tabularx}
    }                                                                                      \\\hline
  \end{tabularx}
  \caption{Surmised misconceptions identified by the SMoL~Tutor. Answers marked
    with ``**'' represent the misconception.}
  \label{t:tutor-surmised-misconceptions}
\end{table}

@@

\begin{figure}[t]

  \includegraphics[width=1\textwidth]{images/Misconceptions_Decay.png}

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{images/smol-quizzes-example.png}
  \caption{Screenshot of a SMoL Quiz question.}
  \label{f:smol-quizzes-example}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{images/smol-tutor-example.png}
  \caption{Screenshots of an interpreting question in SMoL Tutor. The
    top-left shows the initial state, where the question is presented
    as an MCQ. If a student chooses a wrong answer, they will receive
    feedback (bottom-left) and will be asked a similar question
    (right). The similar question must be answered by typing.}
  \label{f:smol-tutor-example}
\end{figure}

@@

\begin{table}[t]
  \centering\begin{tabularx}{\linewidth}{p{0.2\linewidth} X}
    \textbf{Tutorial}                             & \textbf{Goal Sentences}
    \\\hline
    \newterm{scope}: Variable definitions and function definitions &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Referring to an unbound variable leads to an error.
      \item Define "blocks".
      \item SMoL is lexically scoped rather than dynamically scoped.
      \item SMoL disallows defining a variable twice in one block.
      \item SMoL is eager (and evaluates from left to right) rather
            than lazy, reactive, or relational.
    \end{enumerate}
    \\\hline
    \newterm{mut-vars}: Variable updates &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Variable assignment respects the hierarchical structure of blocks.
      \item Variables do not alias.
    \end{enumerate}
    \\\hline
    \newterm{begin}: Sequencing expressions &
    A sequencing expression evaluates its sub-expressions from left to right and
    returns the value of the last sub-expression.
    \\\hline
    \newterm{vectors}: Vectors and vector updates &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Define heap and memory addresses.
      \item Vectors can be referred to by vectors, including themselves.
      \item Vectors are not copied but aliased by bindings.
      \item Constructing vectors doesn't alter environments.
      \item Introducing bindings doesn't alter the (value part of) heap.
    \end{enumerate}
    \\\hline
    \newterm{lambda}: Lambda expressions &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Functions are (first-class) values.
      \item Functions remember the environment where they were created.
      \item Functions can be created with lambda expressions.
      \item A function definition can be viewed as a variable definition plus a lambda expression.
    \end{enumerate}
    \\\hline
    \newterm{local}: Local binding forms &
    Introduce \texttt{let}, \texttt{letrec}, and \texttt{let*}.
    % \\\hline
  \end{tabularx}
  \caption{SMoL Tutor tutorials and their goal sentences.}
  \label{t:tutor-goal-sentences}
\end{table}

@@

  \begin{table}[!h]
    \centering\begin{tabularx}{\linewidth}{X l X}
      \textbf{Misconception}                                                     &
      \textbf{Question}                                                          &
      \textbf{Table of Answers}
      \\\hline
      \miscon{CallByRef} Function calls alias variables.                         &
       % mut-vars.not_aliased_by_funarg_2 (35)
       \lstinputlisting{programs/call-by-ref.txt}                                                                    & {
        \begin{tabularx}{\linewidth}{r X}
          78\%  & \verb|12|      \\
          11\% & \verb|error|      \\
          **10\%   & \verb|0|  \\
          1\%   & \verb|23|  \\
        \end{tabularx}
    }
      \\\hline
      \miscon{CallsCopyStructs} Function calls copy data structures.      &
      % vectors2.alias_with_funcall (48)
      \lstinputlisting{programs/calls-copy-structs.txt}
      & {
        \begin{tabularx}{\linewidth}{r X}
          90\%  & \verb|#(173 0)|      \\
          **10\%   & \verb|#(1 0)|  \\
        \end{tabularx}
    } \\\hline
      \miscon{DeepClosure} Closures copy the \textit{values} of free variables.  &
      % lambda3.lambda_remembers_env (70)
      \lstinputlisting{programs/deep-closure.txt}                                & {
          \begin{tabularx}{\linewidth}{r X}
            86\%  & \verb|4|      \\
            **10\% & \verb|3|      \\
            3\%   & \verb|error|  \\
            1\%   & \verb|lambda| \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{DefByRef} Variable definitions alias variables.                    &
      % mut-vars.not_aliased_by_defvar_1 (29)
      \lstinputlisting{programs/def-by-ref.txt}                                  & {
          \begin{tabularx}{\linewidth}{r X}
            85\%   & \verb|0 12|                \\
            **12\% & \verb|0 0|                 \\
            2\%    & \verb|error|               \\
            1\%    & depends on implementation. \\
          \end{tabularx}
        }
      \\\hline
      \miscon{DefOrSet}  Both definitions and variable assignments are
      interpreted as follows: if a variable is not defined in the
      current environment, it is defined. Otherwise, it is mutated to
      the new value.
      &
      % program: mut-vars.update_undefined (28)
      \lstinputlisting{programs/def-or-set.txt}                                  & {
          \begin{tabularx}{\linewidth}{r X}
            85\%   & \verb|error| \\
            **15\% & \verb|2|     \\
          \end{tabularx}
        }
      \\\hline
      \miscon{DefsCopyStructs} Variable definitions copy structures recursively. &
      % vectors2.alias_with_defvar (47)
      \lstinputlisting{programs/defs-copy-structs.txt}                                & {
          \begin{tabularx}{\linewidth}{r X}
            **67\% & \verb|#(100)| \\
            30\%   & \verb|#(200)| \\
            1\%    & \verb|#(300)| \\
            1\%    & \verb|error|  \\
          \end{tabularx}
      }                                                                                   \\\hline
    \end{tabularx}
    \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
      with ``**'' represent the misconception. (Part I)}
    \label{t:tutor-grounded-misconceptions-1}
  \end{table}

  \begin{table}[!h]

@@

    \begin{tabularx}{\linewidth}{X l X}

@@

          \begin{tabularx}{\linewidth}{r X}

@@

          \begin{tabularx}{\linewidth}{r X}

@@

\begin{table}[!h]
  \centering
  \begin{tabularx}{\linewidth}{X l X}
    \textbf{Misconception}                                                             &
    \textbf{Question}                                                                  &
    \textbf{Table of Answers}                                                              \\\hline
    \surmisedMiscon{NestedDef} Sequences of definitions are
    interpreted as if they are written in nested blocks. A definition
    is not in the scope of later definitions.
    &
    % program: scope.what_is_x_4 (13)
    \lstinputlisting{programs/nested-def.txt}                                          & {
        \begin{tabularx}{\linewidth}{r X}
          92\%  & \verb|2| \\
          **8\% & \verb|1| \\
        \end{tabularx}
    }                                                                                      \\\hline
    \surmisedMiscon{Lazy} Expressions are only evaluated when their values are needed. &
    % order.bad_order (20)
    \lstinputlisting{programs/lazy.txt}                                                & {
        \begin{tabularx}{\linewidth}{r X}
          **57\% & \verb|1 3|   \\
          43\%   & \verb|error| \\
        \end{tabularx}
    }                                                                                      \\\hline
  \end{tabularx}
  \caption{Surmised misconceptions identified by the SMoL~Tutor. Answers marked
    with ``**'' represent the misconception.}
  \label{t:tutor-surmised-misconceptions}
\end{table}

@@

\begin{figure}[!h]

  \includegraphics[width=\linewidth]{images/Misconceptions_Decay.png}
